---
title: "Lab 1"
author: "Markus Lidström"
date: "11/12/2018"
output: pdf_document
---
```{r generate freq table}
tab1 <- as.table(rbind(c(309,191),c(319,281)))
dimnames(tab1) <- list(gender = c("women","men"), opinion = c("favor", "against"))
addmargins(tab1)
addmargins(prop.table(tab1,1),2) # row percentages
addmargins(prop.table(tab1,1),1) # column percentages
prop <- prop.table(tab1)
```
# Exercise 1:1
## 1
```{r}
men_favour <- 319/600
women_favour <- 309/500
men_favour
women_favour
prop
```

For the percentage of men/women in favour, we have 600 men where 319 are in favour so $319/600= men_favour$ and we have 500 women where 309 are in favour so $319/600= women_favour$.

## 2
Let $\theta$ be the odds ratio then we have $$H_0:(\pi_{men},\pi_{women})=(\pi,\pi)$$ $$H_a:(\pi_{men},\pi_{women})\neq(\pi,\pi)$$
Under $H_0$ the proportion of the population in favour of abortion is 628/1100=p and thus under $H_0$ $\mu_{men,in favour}=600p$,$\mu_{women,in favour}=500p$, $\mu_{men,against}=600(1-p)$,$\mu_{women,against}=500(1-p)$ and then we can easily calculate $$G^2 = 2\sum_{j=1}^4n_jlog(n_j/\mu_j)$$ and $$X^2 = \sum_{j=1}^4(n_j-\mu_j)^2/\mu_j$$

## 3

```{r}
## see page 70
309*281/(319*191) ## odds ratio women more in favour
th <- log(309*281/(319*191))
SE <- sqrt(1/309+1/281+1/319+1/191)
ci <- qnorm(c(0.025,0.975))*SE+th
ci
exp(ci)
```

## 4
```{r}
rci <- log(women_favour/men_favour) + qnorm(c(0.025,0.975)) * sqrt((1-men_favour)/600 + (1-women_favour)/500)
exp(rci)
```

```{r}
proportion <- 628/1100
umen <- 600*proportion
uwomen <- 500*proportion
G <- 2*(319*log(319/umen)+
         309*log(309/uwomen)+
         281*log(281/(600-umen))+
         191*log(191/(500-uwomen)))

X <- (319-umen)^2/umen + (309-uwomen)^2/uwomen + (281-(600-umen))^2/(600-umen) + (191-(500-uwomen))^2/(500-uwomen)  
G
X
```



```{r calculate X^2 G^2 p-values} 
chisq.test(tab1,correct=FALSE)
library(MASS)
loglm(~gender+opinion,tab1)
```

```{r calculate risk/odds ratio and CI}
library(epitools)
oddsratio(tab1,rev="neither")
riskratio(tab1,rev="neither")
```

# Exercise 1:2

# Exercise 1:3


